---
title: "Reproducible Research Assignment -1"
---

```{r Packages, echo=TRUE, message = F, warning = F}
    library(knitr)
    library(dplyr)
    library(ggplot2)
```
## Download data set & Reading the data frame

```{r Download the files }
    fileURL<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    download.file(fileURL,destfile="./dataset.zip",method="curl")
    unzip(zipfile = "./dataset.zip")
    activity <- read.csv("./activity.csv",stringsAsFactors = FALSE)
    activity$date <- as.Date(activity$date)
```

## Histogram of Total steps per day

```{r Histogram plot , echo =TRUE, fig.height=6, fig.width=6}

    TotalStepsPerDay <- with(activity,tapply(steps, date, sum))
    hist(TotalStepsPerDay, breaks = 50,col="steelblue",xlab = "Total Steps per Day",
            ylab ="Frequency", main =" Histogram: Total Steps Per Day")
```

##Mean of steps taken each day

```{r  mean , echo= TRUE}
    mean(TotalStepsPerDay,na.rm =TRUE)
```

##Median of steps taken each day
```{r  median, echo= TRUE}
    median(TotalStepsPerDay,na.rm =TRUE)
```

##Time series plot of the average number of steps taken
```{r  plot, echo= TRUE}

    steps_by_interval <- aggregate(steps ~ interval, activity, mean)

    g  <- ggplot(steps_by_interval, aes(interval, steps), colour = 'red') + geom_line()  
    g + xlab("5-minute interval") + ylab("avarage number of steps")
   
```

## The 5-minute interval that, on average, contains the maximum number of steps
 - Time Series that contain activity of maximum steps 
```{r Max steps, echo= TRUE}
    max_steps<- max(steps_by_interval$steps)
    yy <- which(grepl(max_steps, steps_by_interval$steps))
    steps_by_interval[yy,]
```

 - Average no of steps  
```{r avg steps, echo= TRUE}
    avg_steps <- mean(steps_by_interval$steps)
    avg_steps
```
- Total no of Time Series that contain steps above average
```{r Max sum of avg steps, echo= TRUE}
    sum(steps_by_interval$steps>=avg_steps )
```


## Code to describe and show a strategy for imputing missing data

```{r No activity, echo= TRUE }
    No_activity<-is.na(activity$steps)
    sum(No_activity)
```


```{r impute, echo = TRUE}

for (i in 1:nrow(activity)) {
    if(is.na(activity$steps[i])) {
        val <- steps_by_interval$steps[which(steps_by_interval$interval == activity$interval[i])]
        activity$steps[i] <- val 
    }
}

# Aggregate the steps per day with the imputed values
impute_steps_per_day <- aggregate(steps ~ date, activity, sum)

# Draw a histogram of the value 
hist(impute_steps_per_day$steps, main = "Histogram of total number of steps per day (Imputed)", xlab = "Steps per day")
```

##Are there differences in activity patterns between weekdays and weekends?
- Subseting the data set by weekdays and weekends
- Convering the day type column to a factor column
- Aggregate the Weekday and weekend activity 

```{r  Subsetting the data , echo= TRUE}
    activity['day_type'] <-weekdays(activity$date)
    activity$day_type[activity$day_type  %in% c('Saturday','Sunday') ] <- "weekend"
    activity$day_type[activity$day_type != "weekend"  ] <- "weekday" 
    activity$day_type <- as.factor(activity$day_type)
    stepsMeanPerWeekday <- aggregate(steps ~ interval + day_type, activity, mean)
```
- Plot of activity patterns between weekdays and weekends?

```{r ploting the ggraph, echo= TRUE}
# Display the 2 plots
    qplot(interval,  steps, data = stepsMeanPerWeekday, geom=c("line"), xlab = "Interval", 
          ylab ="Number of steps", color = day_type) +facet_wrap(~ day_type, ncol = 1)
```

    
